An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-11-29 paper/1.introduction.tex
add explanation for breverse and titlerunning

Here, let \texttt{[1,2]} be the original source \texttt{s}.
$get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
%\texttt{get pHead [1,2] = 1}.
After the view \texttt{v} is obtained, the user can modify the view.
In this case, the user
modified the view
from \texttt{1} to \texttt{100}.
From this updated view, how can the source be updated?
This is ``the view update problem.''
To solve this problem, BX provides $put$, an update function on the original source.
In this example, $put$ of $phead$ constructs a new list \texttt{[100,2]} from the updated view (\texttt{v'}) and the original source (\texttt{s}).
Note that the update to s' uses s and v' as input.
%\texttt{put pHead [1,2] 8 = [8,2]}

@@

Figure \ref{fig:eval-comp-phead} illustrates the evaluation of these programs where the original source \texttt{s} is \texttt{[[[1,2],[3]],[4]]} and the updated view is \texttt{100}.
By repetitious evaluation of $get$s of $phead$, we can obtain the view $v$, \texttt{1}. To obtain the final updated source \texttt{s'}, we need to evaluate $put$ three times. The first put evaluation is from \texttt{i$_2$} and \texttt{v'} and we can obtain \texttt{i$_2$'}.

@@

If the programs are simple like these running examples, we can transform left associative compositions to right associative. However, this is not always possible. Therefore it is important to have a fast evaluation method for left associative compositions.
For achieving this, we introduce two methods that use memoization.
The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed.'' There is no quadratic $get$s in this method, therefore this method improves the runtime of many BX programs. However, a disadvantage of this is the familiar tradeoff for keeping and searching values in a table. Especially, when we use large inputs, the evaluation time will be long. We explain this method in Section 3.

@@

For obtaining complements, we use tupling of $put$ and $get$, and produce new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
As an example, let us consider the previous example again: $phead \circ (phead \circ phead)$ (rp3) in figure \ref{fig:eval-comp-phead-2}. \texttt{c$_1$} and \texttt{c$_2$} are complements. The points in this figure is two. First, after evaluation of the first $pg$, we do not need to keep the original source \texttt{s}, because all its information is in \texttt{c$_1$} and \texttt{i$_1$}. Second, complements are smaller than the intermediate states in the previous figure.
%This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources. 
Actually, the simple combined $pg$ is not effective for left associative compositions, because this requires other two $put$ in the right part of the figure. To achieve efficient evaluation, we use two techniques, lazy update and lazy evaluation. We explain the second method and optimization in Section 4.

>>>>

Here, let $[1,2]$ be the original source \texttt{s}.
$get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
%\texttt{get pHead [1,2] = 1}.
After the view \texttt{v} is obtained, the user can modify the view.
In this case, the user
modified the view
from $1$ to $100$.
From this updated view, how can the source be updated?
This is ``the view update problem.''
To solve this problem, BX provides $put$, an update function on the original source.
In this example, $put$ of $phead$ constructs a new list ${[100,2]}$ from the updated view (\texttt{v'}) and the original source (\texttt{s}).
Note that the update to s' uses s and v' as input.
%\texttt{put pHead [1,2] 8 = [8,2]}

@@

Figure \ref{fig:eval-comp-phead} illustrates the evaluation of these programs where the original source \texttt{s} is ${[[[1,2],[3]],[4]]}$ and the updated view is ${100}$.
By repetitious evaluation of $get$s of $phead$, we can obtain the view $v$, ${1}$. To obtain the final updated source \texttt{s'}, we need to evaluate $put$ three times. The first put evaluation is from \texttt{i$_2$} and \texttt{v'} and we can obtain \texttt{i$_2$'}.

@@

If the programs are simple like these running examples, we can transform left associative compositions to right associative. However, this is not always possible. \textcolor{red}{For example, let us consider the following 'breverse' program.}

%$\tab
%bsnoc = Rearr\\
%bfoldr bx = Case \ f_1 \ f_2 \ bx_s \text{where: } f_1 = , f_2% = \\

%breverse = RearrS \ f_1 \ f_2 \ bx_s \ \text{where: } f_1 = \l%ambda s.(s.()), \, f_2 = \lambda (s.()).s,\\
%\qtab bx_s = bfoldr bsnoc$

\textcolor{red}{In this program, recursion occurs on the left-hand side of composition. Because of priority, it is impossible to tranform this program to right associative compositions. At the same time, we can not apply fusion \cite{} to this program because this program is dynamically produced depending on the length of the input list.}
Therefore it is important to have a fast evaluation method for left associative compositions.

For achieving this, we introduce two methods that use memoization.
The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed.'' There is no quadratic $get$s in this method, therefore this method improves the runtime of many BX programs. However, a disadvantage of this is the familiar tradeoff for keeping and searching values in a table. Especially, when we use large inputs, the evaluation time will be long. We explain this method in Section 3.

@@

For obtaining complements, we use tupling of $put$ and $get$, and produce new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
As an example, let us consider the previous example again: $phead \circ phead \circ phead$ in figure \ref{fig:eval-comp-phead-2}. \texttt{c$_1$} and \texttt{c$_2$} are complements. \texttt{d$_1$} and \texttt{d$_2$} are valid views for \texttt{s} and \texttt{i$_1$}. The points in this figure is two. First, after evaluation of the first $pg$, we do not need to keep the original source \texttt{s}, because all its information is in \texttt{c$_1$} and \texttt{i$_1$}. Second, complements are smaller than the intermediate states in the previous figure.
%This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources. 
Actually, the simple combined $pg$ is not effective for left associative compositions, because this requires other two $put$ in the right part of the figure. To achieve efficient evaluation, we use two techniques, lazy update and lazy evaluation. We explain the second method and optimization in Section 4.
